### 2015-02-04 cav-2015-tex/approach.tex
pass through section 2.2

\subsection{Generating Summary Heaps}
\label{sec:precise}
The summary
heap algorithm only
partitions the heap at points of divergence in the control flow graph
such as field accesses or writes (execeptional control flow due to null
pointers) or comparisons of references (program control flow). There
are three sets of rewrite rules specific to the summary heap
algorithm: (i) rules to initialize symbolic references, (ii) rules to
complete field access and field write, and (iii) rules to check
equality and inequality of references.

@@

\begin{figure*}
\begin{tabular}[c]{l}
\scalebox{1.0}{\usebox{\boxPI}} \\
\end{tabular}
\caption{Summary machine, $s ::= \lp\cfgnt{L}\ \cfgnt{R}\ \cfgnt{r}\ \cfgnt{f}\ \cfgnt{C}\rp$, with $s\rsum^*s^\prime =  s \rsum \cdots \rsum s^\prime \rsum s^\prime$.}
\label{fig:symInit}
\end{figure*}

The initialization rule in \figref{fig:symInit} is invoked whenever a
field pointing to an uninitialized reference is accessed. The
interaction with the solver in the definition of the rules is denoted
by $\mathbb{S}(\phi)$ which returns true if $\phi$ is satisfiable.
%The summarize-end rule becomes active when there is nothing to
%initialize.
 The check for uninitialized references is determined by the function
 $\mathbb{UN}(\cfgnt{L}, \cfgnt{R}, \cfgnt{r}, \cfgnt{f}) =
 \{\lp\phi\ \cfgnt{l}\rp\ ...\}$ which returns constraint-location
 pairs where the field $\cfgnt{f}$ is uninitialized.
\[
\begin{array}{rcl}
\mathbb{UN}(\cfgnt{L}, \cfgnt{R}, \cfgnt{r}, \cfgnt{f}) & = &\{ \lp\phi\ \cfgnt{l}\rp \mid \lp \phi\ \cfgnt{l}\rp  \in \cfgnt{L}\lp \cfgnt{r}\rp  \wedge \\
& & \ \ \ \ \exists \phi^\prime \lp \lp \phi^\prime\ \cfgnt{l}_\mathit{un}\rp  \in \cfgnt{L}\lp \cfgnt{R}\lp l,\cfgnt{f}\rp\rp \wedge \\
& & \ \ \ \ \ \ \ \ \mathbb{S}\lp \phi \wedge \phi^\prime \rp\rp\}\\
\end{array}
\]

In~\figref{fig:symInit}, given the uninitialized set $\Lambda$ for
field $f$, the $\mathit{min}_l$ function returns
$(\phi_x\ \cfgnt{l}_x)$ based on a lexicographical ordering of
uninitialized locations in $\Lambda$ to make initialization
deterministic.  The set $\rho$ contains the set of reference-location
pairs from the input heap that are potential aliases.  There are four
choices: (i) $\theta_\mathit{null}$ asserts the condition under which
$\cfgnt{l}_\mathit{null}$ is possible, (ii) $\theta_\mathit{new}$ is
active when $\cfgnt{r}_f$ points to the fresh location, (iii) each
member of $\theta_\mathit{alias}$ restricts $\cfgnt{r}_f$ to a
particular alias in $\rho$, and at the same time, not alias any member
of $\rho$ that was initialized earlier than the current choice, and
(iv) $\theta_\mathit{orig}$ implements conditional initialization to
preserve the original heap structure.  These sets are added into the
heap on $\cfgnt{r}_f$ after the fields for $\cfgnt{l}_f$ are
initialized to $\cfgnt{r}_\mathit{un}$.

@@

We visualize the initialization process in~\figref{fig:initHeap}. The
graph in~\figref{fig:initHeap}(a) represents the initial heap. The
reference superscripts $s$ and $i$ indicate the partition containing
the reference.  In~\figref{fig:initHeap}, $\cfgnt{r}_0^s$ represents a
stack reference for the $\cfgt{this}$ variable which has two fields
$x$ and $y$ of the same type. Note that when no constraint is
specified for a location, there is an implicit $\mathit{true}$
constraint. For example, $\cfgnt{r}_0^s$ points to $\cfgnt{l}_0$ on
the constraint $\mathit{true}$. The fields $x$ and $y$ point to the
uninitialized reference $\cfgnt{r}_\mathit{un}$. The graph
in~\figref{fig:initHeap}(b) represents the summary heap after the
initialization of the $\lp\cfgt{this}\ \cfgt{\$}\ \cfgnt{x}\rp$ field
while the graph in~\figref{fig:initHeap}(c) represents the symbolic
heap summary after the initialization of the
$\lp\cfgt{this}\ \cfgt{\$}\ \cfgnt{y}\rp$ field following the
initialization of $\lp\cfgt{this}\ \cfgt{\$}\ \cfgnt{x}\rp$. The list
in~\figref{fig:initHeap}(d) represents the various sets constructed in
the summarize rule during the initialization of
$\lp\cfgt{this}\ \cfgt{\$}\ \cfgnt{y}\rp$. We also define values of
constraints used as labels in the graph.

@@

There are two rewrite rules in~\figref{fig:fHeap}(a), one for reading
the value of a field (field-access) and the other when we write to a
field (field-write). Both rules first check that the operations can be
performed on a non-null location.  In the field-access rewrite rule
in~\figref{fig:fHeap} gets the field type $\cfgnt{C}$ and uses the
summarize machine, \figref{fig:symInit}, to get a new summary heap
that is initialized on the field $\cfgnt{f}$. After initialization,
the summary heap is further modified to include a new stack reference
pointing to the value set from the read. The new stack reference is
the return value from the operation.

@@

The value set computation is local: it only considers constraint
locations on the reference $\cfgnt{r}$. Further, it only includes
things that are feasible under the current path condition $\phi_g$. It
also ensures that the access path is valid. There is a constraint
$\phi$ from the base reference $r$ and a given location. Once a
location $l$ is selected, the field $\cfgnt{f}$ is accessed to get
another reference $r^\prime$, and from that reference there is a set
of locations each with a constraint. The first constraint $\phi$ from
the base reference plus this second constraint to the actual location
of the field $\phi^\prime$ comprise the access path. For each member
of the value set, these constraints together with the global path
constraint must be satisfied on some aliasing assignment.

@@

The rewrite rules to compare two references in the symbolic summary
heap are shown in~\figref{fig:eqs}. The equals references-true rewrite
rule returns true when two references $\cfgnt{r}_0$ and $\cfgnt{r}_1$
\emph{can} be equal. Intutively, $\Phi_\alpha$ contains all
constraints under which $\cfgnt{r}_0$ and $\cfgnt{r}_1$ may point to
the same location in the symbolic summary heap. The second set,
$\Phi_0$, contains constraints under which the reference $\cfgnt{r}_0$
points to corresponding locations such that the reference
$\cfgnt{r}_1$ \emph{does not} point to those locations under any
constraint. Finally, the set, $\Phi_1$, contains constraints under
which $\cfgnt{r}_1$ points to corresponding locations and
$\cfgnt{r}_0$ \emph{does not} point to those locations under any
constraint. We use the three sets of constraints to update the current
global heap constraint $\phi_g$ and create a new global heap
constraint $\phi_g^\prime$. We add to $\phi_g$ the disjunction of the
constraints in $\Phi_\alpha$ to indicate that if any of the
constraints are satisfiable, then references $\cfgnt{r}_0$ and
$\cfgnt{r}_1$ can be equal. Furthermore, we conjoin to the global heap
constraint $\phi_g^\prime$, the conjunctions of negations to the
constraints in $\Phi_0$ and $\Phi_1$. This indicates for locations
that are not common to the references, the negations of their
constraints are satisfiable. The rule does not complete (i.e., is not
feasible) if the new global constraint is not satisfied on any
aliasing assignment. In such a case, the $\cfgt{true}$ outcome is not
possible on any heap in the summary. Before the rewrite rule returns
$\cfgt{true}$ we check the satisfiability of the updated global heap
constraint. The reference-false is the logical dual of the rule.

>>>>

\subsection{Generating Symbolic Heaps}
\label{sec:precise}

The symbolic heap algorithm only partitions the heap at points of
divergence in the control flow graph such as comparisons of references
(program control flow) and field dereferences or writes (execeptional
control flow due to null pointers). There are three sets of rewrite
rules specific to the symbolic heap algorithm: (i) rules to initialize
symbolic references, (ii) rules to perform field dereferences and
writes, and (iii) rules to check equality and inequality of
references.

@@

\begin{figure*}
\begin{tabular}[c]{l}
\scalebox{1.0}{\usebox{\boxPI}} \\
\end{tabular}
\caption{Initializing fields, $s ::= \lp\cfgnt{L}\ \cfgnt{R}\ \cfgnt{r}\ \cfgnt{f}\ \cfgnt{C}\rp$, with $s\rsum^*s^\prime =  s \rsum \cdots \rsum s^\prime \rsum s^\prime$.}
\label{fig:symInit}
\end{figure*}

The initialization rule in \figref{fig:symInit} is invoked whenever a
field within an uninitialized reference is accessed. The interaction
with the solver in the definition of the rules is denoted by
$\mathbb{S}(\phi)$ which returns true if $\phi$ is satisfiable.  The
check for uninitialized references is determined by the function
$\mathbb{UN}(\cfgnt{L}, \cfgnt{R}, \cfgnt{r}, \cfgnt{f}) =
\{\lp\phi\ \cfgnt{l}\rp\ ...\}$ which returns constraint-location
pairs where the field $\cfgnt{f}$ is uninitialized.
\[
\begin{array}{rcl}
\mathbb{UN}(\cfgnt{L}, \cfgnt{R}, \cfgnt{r}, \cfgnt{f}) & = &\{ \lp\phi\ \cfgnt{l}\rp \mid \lp \phi\ \cfgnt{l}\rp  \in \cfgnt{L}\lp \cfgnt{r}\rp  \wedge \exists \phi^\prime \lp \lp \phi^\prime\ \cfgnt{l}_\mathit{un}\rp  \in \cfgnt{L}\lp \cfgnt{R}\lp l,\cfgnt{f}\rp\rp \wedge \mathbb{S}\lp \phi \wedge \phi^\prime \rp\rp\}\\
\end{array}
\]

In~\figref{fig:symInit}, given the uninitialized set $\Lambda$ for
field $f$, the $\mathit{min}_l$ function returns
$(\phi_x\ \cfgnt{l}_x)$ based on a lexicographical ordering of
uninitialized locations in $\Lambda$ to make initialization
deterministic.  The set $\rho$ contains reference-location pairs that
represent potential aliases. There are four cases encoded in the
symbolic heap: (i) $\theta_\mathit{null}$ represents the condition
where $\cfgnt{l}_\mathit{null}$ is possible, (ii)
$\theta_\mathit{new}$ represents the case where $\cfgnt{r}_f$ points
to a fresh location, (iii) each member of $\theta_\mathit{alias}$
restricts $\cfgnt{r}_f$ to a particular alias in $\rho$, and at the
same time, not alias any member of $\rho$ that was initialized earlier
than the current choice, and (iv) $\theta_\mathit{orig}$ implements
conditional initialization to preserve the original heap structure.
These sets are added into the heap on $\cfgnt{r}_f$ after the fields
for $\cfgnt{l}_f$ are initialized to $\cfgnt{r}_\mathit{un}$.

@@

We visualize the initialization process in~\figref{fig:initHeap}. The
graph in~\figref{fig:initHeap}(a) represents the initial heap. The
reference superscripts $s$ and $i$ indicate the partition containing
the reference.  In~\figref{fig:initHeap}, $\cfgnt{r}_0^s$ represents a
stack reference for the $\cfgt{this}$ variable which has two fields
$x$ and $y$ of the same type. Note that when no constraint is
specified for a location, there is an implicit $\mathit{true}$
constraint. For example, $\cfgnt{r}_0^s$ points to $\cfgnt{l}_0$ on
the constraint $\mathit{true}$. The fields $x$ and $y$ point to the
uninitialized reference $\cfgnt{r}_\mathit{un}$. The graph
in~\figref{fig:initHeap}(b) represents the symbolic heap after the
initialization of the $\lp\cfgt{this}\ \cfgt{\$}\ \cfgnt{x}\rp$ field
while the graph in~\figref{fig:initHeap}(c) represents the symbolic
heap after the initialization of the
$\lp\cfgt{this}\ \cfgt{\$}\ \cfgnt{y}\rp$ field following the
initialization of $\lp\cfgt{this}\ \cfgt{\$}\ \cfgnt{x}\rp$. The list
in~\figref{fig:initHeap}(d) represents the various sets constructed in
initialization rule for $\lp\cfgt{this}\ \cfgt{\$}\ \cfgnt{y}\rp$. We
also define values of constraints used as labels in the graph.

@@

There are two rewrite rules in~\figref{fig:fHeap}(a), one for reading
the value of a field (field-access) and the other when we write to a
field (field-write). Both rules first check that the operations can be
performed on a non-null location.  In the field-access rewrite rule
in~\figref{fig:fHeap} gets the field type $\cfgnt{C}$ and uses
\figref{fig:symInit}, to get a new symbolic heap that is initialized
on the field $\cfgnt{f}$. The symbolic heap is further modified to
include a new stack reference pointing to the value set from the read;
the new stack reference is the return value from the operation.

@@

The value set computation is local: it only considers constraint
locations on the reference $\cfgnt{r}$. Further, it only includes
constraints that are feasible under the current path condition
$\phi_g$. It also ensures that the access path is valid. There is a
constraint $\phi$ from the base reference $r$ and a given
location. Once a location $l$ is selected its field $\cfgnt{f}$ is
accessed to get another reference $r^\prime$, and from that reference
there is a set of locations each with a constraint. The first
constraint $\phi$ from the base reference plus this second constraint
to the actual location of the field $\phi^\prime$ comprise the access
path. For each member of the value set, these constraints together
with the global path constraint must be satisfied on some aliasing
assignment. \nsr{neha: this para needs work -- flow and content}

@@

The rewrite rules to compare two references in the symbolic heap are
shown in~\figref{fig:eqs}. The equals references-true rewrite rule
returns true when two references $\cfgnt{r}_0$ and $\cfgnt{r}_1$
\emph{can} be equal. Intutively, $\Phi_\alpha$ contains all
constraints under which $\cfgnt{r}_0$ and $\cfgnt{r}_1$ may point to
the same location in the symbolic heap. The second set, $\Phi_0$,
contains constraints under which the reference $\cfgnt{r}_0$ points to
corresponding locations such that the reference $\cfgnt{r}_1$
\emph{does not} point to those locations under any
constraint. Finally, the set, $\Phi_1$, contains constraints under
which $\cfgnt{r}_1$ points to corresponding locations and
$\cfgnt{r}_0$ \emph{does not} point to those locations under any
constraint. We use the three sets of constraints to update the current
global heap constraint $\phi_g$ and create a new global heap
constraint $\phi_g^\prime$. We add to $\phi_g$ the disjunction of the
constraints in $\Phi_\alpha$ to indicate that if any of the
constraints are satisfiable, then references $\cfgnt{r}_0$ and
$\cfgnt{r}_1$ can be equal. Furthermore, we conjoin to the global heap
constraint $\phi_g^\prime$, the conjunctions of negations to the
constraints in $\Phi_0$ and $\Phi_1$. This indicates for locations
that are not common to the references, the negations of their
constraints are satisfiable. The rule does not complete (i.e., is not
feasible) if the new global constraint is not satisfied on any
aliasing assignment. In such a case, the $\cfgt{true}$ outcome is not
possible on any symbolic heap. Before the rewrite rule returns
$\cfgt{true}$ we check the satisfiability of the updated global heap
constraint. The reference-false is the logical dual of the rule.
